---
title: "NFL behaviour"
author: "Bindoff, A., Tian. N."
output:
  html_document:
    code_folding: 'hide'
---

`r Sys.time()`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## {.tabset}

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(knitr)
library(reshape2)
library(ggbeeswarm)
library(openxlsx)


# d <- read_csv("nfl_exercise.csv") %>%
#   mutate(exercise = factor(exercise, levels = c(1, 0), labels = c("exercise", "no exercise")),
#          geno = factor(geno, levels = c("WT", "NFL")))

```

```{r}
d0 <- lapply(1:3, function(x) read.xlsx('nfl_behaviour_2019.xlsx', sheet = x, startRow = 3))

names(d0[[3]]) <- c("ID",
                    "Geno",
                    "Exercise",
                    "HLS_pre",
                    "HLS_post",
                    "HLS_change",
                    "FLS_pre",
                    "FLS_post",
                    "FLS_change",
                    "ALS_pre",
                    "ALS_post",
                    "ALS_change")

d <- left_join(d0[[1]], d0[[2]]) %>%
  left_join(d0[[3]]) %>%
  mutate(Exercise = factor(Exercise, levels = c(0, 1), labels = c("No exercise", "Exercise")),
         Geno = factor(Geno, levels = c("wt", "NFL")))

```


### Open field

#### Centre time

```{r}
dw = 2/3

ggplot(d, aes(x = Geno, colour = Exercise, y = CT_change)) +
  geom_boxplot(position = position_dodge(width = dw), alpha = 0) +
  geom_quasirandom(dodge.width = dw, size = 2) +
  scale_colour_manual(values = c("black", "red")) +
  ylab("Change time in centre (%)") +
  xlab("Genotype") +
  theme_bw()
```
##### Unadjusted means

```{r}
d %>% group_by(Geno) %>%
  summarise(mean = mean(CT_change, na.rm = TRUE),
            sd = sd(CT_change, na.rm = TRUE)) %>%
  kable(digits = 2)

d %>% group_by(Exercise) %>%
  summarise(mean = mean(CT_change, na.rm = TRUE),
            sd = sd(CT_change, na.rm = TRUE)) %>%
  kable(digits = 2)

d %>% group_by(Geno, Exercise) %>%
  summarise(mean = mean(CT_change, na.rm = TRUE),
            sd = sd(CT_change, na.rm = TRUE)) %>%
  kable(digits = 2)
```


##### ANCOVA

CT post adjusted for CT pre   

```{r}
m1 <- lm(CT_post ~ CT_pre + Geno * Exercise, d)
anova(m1)

#plot(m1)
```

##### Adjusted means and 95% CI

```{r}
dd <- expand.grid(CT_pre = mean(d$CT_pre),
                   Geno = factor(levels(d$Geno), levels = c("wt", "NFL")),
                   Exercise = factor(levels(d$Exercise), levels = c("No exercise", "Exercise")))

p <- predict(m1, dd, se.fit = TRUE)
dd$lwr <- p$fit - 1.96*p$se.fit
dd$upr <- p$fit + 1.96*p$se.fit
dd$fit <- p$fit

knitr::kable(dd, digits = 2)
```

```{r}
ggplot(dd, aes(x = Geno, colour = Exercise, y = fit, ymin = lwr, ymax = upr)) +
  ggbeeswarm::geom_quasirandom(data = d, dodge.width = 0.4, alpha = 0.2, inherit.aes = FALSE,
             aes(y = CT_post, x= Geno, colour = Exercise)) +
  geom_point(size = 3, position = position_dodge(width = 0.4)) +
  geom_errorbar(position = position_dodge(width = 0.4), size = 1, width = 0.2) +
  scale_colour_manual(values = c("black", "red")) +
  ylab("Adjusted centre time (s)") +
  xlab("") +
  theme_bw()

```



#### Total movement

```{r}
ggplot(d, aes(x = Geno, colour = Exercise, y = TM_change)) +
  geom_boxplot(position = position_dodge(width = dw), alpha = 0.0) +
  geom_quasirandom(dodge.width = dw, size = 2) +
  scale_colour_manual(values = c("black", "red")) +
  ylab("Change total movement (cm)") +
  xlab("Genotype") +
  theme_bw()
```

##### Unadjusted means (total movement change)  

```{r}
d %>% group_by(Geno) %>%
  summarise(mean = mean(TM_change, na.rm = TRUE),
            sd = sd(TM_change, na.rm = TRUE)) %>%
  kable(digits = 2)

d %>% group_by(Exercise) %>%
  summarise(mean = mean(TM_change, na.rm = TRUE),
            sd = sd(TM_change, na.rm = TRUE)) %>%
  kable(digits = 2)

d %>% group_by(Geno, Exercise) %>%
  summarise(mean = mean(TM_change, na.rm = TRUE),
            sd = sd(TM_change, na.rm = TRUE)) %>%
  kable(digits = 2)



```


##### ANCOVA (total movement post adjusted for total movement pre)

```{r}
m1 <- lm(TM_post ~ TM_pre + Geno*Exercise, d)
anova(m1)

#plot(m1)
```

##### Adjusted means and 95% CI

```{r}
dd <- expand.grid(TM_pre = mean(d$TM_pre),
                   Geno = factor(levels(d$Geno), levels = c("wt", "NFL")),
                   Exercise = factor(levels(d$Exercise), levels = c("No exercise", "Exercise")))

p <- predict(m1, dd, se.fit = TRUE)
dd$lwr <- p$fit - 1.96*p$se.fit
dd$upr <- p$fit + 1.96*p$se.fit
dd$fit <- p$fit

knitr::kable(dd, digits = 2)
```

```{r}
ggplot(dd, aes(x = Geno, colour = Exercise, y = fit, ymin = lwr, ymax = upr)) +
  ggbeeswarm::geom_quasirandom(data = d, dodge.width = 0.4, alpha = 0.2, inherit.aes = FALSE,
             aes(y = TM_post, x= Geno, colour = Exercise)) +
  geom_point(size = 3, position = position_dodge(width = 0.4)) +
  geom_errorbar(position = position_dodge(width = 0.4), size = 1, width = 0.2) +
  scale_colour_manual(values = c("black", "red")) +
  ylab("Total movement (cm)") +
  xlab("") +
  theme_bw()

```


### Grip strength

```{r}
gs <- select(d, ID, Geno, Exercise, HS_pre, HS_post, FS_pre, FS_post) %>%
  melt(id.vars = c("ID", "Geno", "Exercise")) %>%
  mutate(time = case_when(variable %in% c("HS_pre", "FS_pre") ~ "Pre",
                          variable %in% c("HS_post", "FS_post") ~ "Post"),
         time = factor(time, levels = c("Pre", "Post"))) %>%
  mutate(Limb = case_when(variable %in% c("HS_pre", "HS_post") ~ "Hindlimb",
                          variable %in% c("FS_pre", "FS_post") ~ "Forelimb"))


ggplot(gs, aes(x = time, colour = Exercise, y = value)) +
  geom_boxplot(position = position_dodge(width = dw), alpha = 0) +
    geom_quasirandom(dodge.width = dw, size = 2, alpha = 1/3, aes(shape = Limb, group = Exercise)) +
  scale_colour_manual(values = c("black", "red")) +
  facet_wrap(~Geno) +
  ylab("Grip strength") +
  xlab("Genotype") +
  theme_bw()

```

Mixed effects models were fitted with a random intercept for animal id using the `lme4` package in R.

```{r eval = FALSE}
m1 <- lmer(value ~ Limb + Geno*time + Exercise*time + Geno*Exercise*time + (1|ID), gs)
anova(m1)

gs %>% group_by(Geno, time) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE)) %>%
  kable(digits = 2)

gs %>% group_by(Exercise, time) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE)) %>%
  kable(digits = 2)
```


```{r eval = FALSE}
m2 <- lmer(value ~ Geno*Exercise*time + (1|ID), gs)
gs0 <- expand.grid(time = factor(levels(gs$time), levels = c("Pre", "Post")),
                   Geno = factor(levels(gs$Geno), levels = c("wt", "NFL")),
                   Exercise = factor(levels(gs$Exercise), levels = c("No exercise", "Exercise")))

bootfit <- bootMer(m2, FUN=function(x) predict(x, gs0, re.form = NA),
                   nsim = 500,
                   parallel = "multicore",
                   ncpus = 3L)

gs0$lwr <- apply(bootfit$t, 2, quantile, 0.025)
gs0$upr <- apply(bootfit$t, 2, quantile, 0.975)
gs0$fit <- apply(bootfit$t, 2, mean)


ggplot(gs0, aes(x = time, colour = Geno, group = Geno, y = fit, ymin = lwr, ymax = upr)) +
  geom_point(size = 3, position = position_dodge(width = 0.2)) +
  geom_errorbar(position = position_dodge(width = 0.2), size = 1, width = 0.2) +
  geom_path(position = position_dodge(width = 0.2), linetype = "dashed", size = 1) +
  scale_colour_manual(values = c("black", "red")) +
  ylab("Grip strength /gm bw") +
  xlab("") +
  facet_wrap(~Exercise) +
  ylim(40, 90) +
  theme_bw()

```

```{r eval = FALSE}
ggplot(gs0, aes(x = time, colour = Exercise, group = Exercise, y = fit, ymin = lwr, ymax = upr)) +
  geom_point(size = 3, position = position_dodge(width = 0.2)) +
  geom_errorbar(position = position_dodge(width = 0.2), size = 1, width = 0.2) +
  geom_path(position = position_dodge(width = 0.2), linetype = "dashed", size = 1) +
  scale_colour_manual(values = c("black", "red")) +
  ylab("Grip strength /gm bw") +
  xlab("") +
  facet_wrap(~Geno) +
  ylim(40, 90) +
  theme_bw()

```

##### Unadjusted means (change in grip-strength)

```{r}
gs1 <- gs %>% select(ID, Geno, Exercise, time, value, Limb) %>% tidyr::spread(time, value)


gs1 %>% group_by(Geno, Limb) %>%
  summarise(gs.mean = mean(Post - Pre),
            gs.sd = sd(Post - Pre)) %>%
  knitr::kable(digits = 2)

gs1 %>% group_by(Exercise, Limb) %>%
  summarise(gs.mean = mean(Post - Pre),
            gs.sd = sd(Post - Pre)) %>%
  knitr::kable(digits = 2)

gs1 %>% group_by(Geno, Exercise, Limb) %>%
  summarise(gs.mean = mean(Post - Pre),
            gs.sd = sd(Post - Pre)) %>%
  knitr::kable(digits = 2)
```

##### Mixed ANCOVA

In order to account for measurements of both front- and hind-limb, a mixed model ANCOVA was fitted with a random intercept for mouse ID.  

```{r}
m1 <- lmer(Post ~ Pre + Geno*Exercise + (1|ID), gs1)
anova(m1)
```

```{r, fig.cap = "Mean grip strength [95% CI]"}

gs0 <- expand.grid(Pre = mean(gs1$Pre),
                   Geno = factor(levels(gs1$Geno), levels = c("wt", "NFL")),
                   Exercise = factor(levels(gs1$Exercise), levels = c("No exercise", "Exercise")))

bootfit <- bootMer(m1, FUN=function(x) predict(x, gs0, re.form = NA),
                   nsim = 500,
                   parallel = "multicore",
                   ncpus = 3L)

gs0$lwr <- apply(bootfit$t, 2, quantile, 0.025)
gs0$upr <- apply(bootfit$t, 2, quantile, 0.975)
gs0$fit <- apply(bootfit$t, 2, mean)

knitr::kable(gs0, digits = 2)

```

```{r}

ggplot(gs0, aes(x = Exercise, colour = Geno, group = Geno, y = fit, ymin = lwr, ymax = upr)) +
  ggbeeswarm::geom_quasirandom(data = gs1, dodge.width = 0.4, alpha = 0.2, inherit.aes = FALSE,
             aes(y = Post, x= Exercise, colour = Geno)) +
  geom_point(size = 3, position = position_dodge(width = 0.4)) +
  geom_errorbar(position = position_dodge(width = 0.4), size = 1, width = 0.2) +
  #geom_path(position = position_dodge(width = 0.2), linetype = "dashed", size = 1) +
  scale_colour_manual(values = c("black", "red")) +
  ylab("Grip strength gm/gm bw") +
  xlab("") +
  ylim(40, 100) +
  theme_bw()

```

