---
title: "NFL behaviour"
author: "Bindoff, A., Tian. N."
output:
  html_document:
    code_folding: 'hide'
---

`r Sys.time()`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## {.tabset}

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(knitr)
library(reshape2)
library(ggbeeswarm)
library(openxlsx)


# d <- read_csv("nfl_exercise.csv") %>%
#   mutate(exercise = factor(exercise, levels = c(1, 0), labels = c("exercise", "no exercise")),
#          geno = factor(geno, levels = c("WT", "NFL")))

```

```{r}
d0 <- lapply(1:3, function(x) read.xlsx('nfl_behaviour_2019.xlsx', sheet = x, startRow = 3))

names(d0[[3]]) <- c("ID",
                    "Geno",
                    "Exercise",
                    "HLS_pre",
                    "HLS_post",
                    "HLS_change",
                    "FLS_pre",
                    "FLS_post",
                    "FLS_change",
                    "ALS_pre",
                    "ALS_post",
                    "ALS_change")

d <- left_join(d0[[1]], d0[[2]]) %>%
  left_join(d0[[3]]) %>%
  mutate(Exercise = factor(Exercise, levels = c(0, 1), labels = c("No exercise", "Exercise")),
         Geno = factor(Geno, levels = c("wt", "NFL")))

```


### Open field

#### Centre time

```{r}
dw = 2/3

ggplot(d, aes(x = Geno, colour = Exercise, y = CT_change)) +
  geom_boxplot(position = position_dodge(width = dw), alpha = 0) +
  geom_quasirandom(dodge.width = dw, size = 2) +
  scale_colour_manual(values = c("black", "red")) +
  ylab("Change time in centre (%)") +
  xlab("Genotype") +
  theme_bw()
```

#### ANCOVA

CT post adjusted for CT pre   

```{r}
m1 <- lm(CT_post ~ CT_pre + Geno * Exercise, d)
anova(m1)

#plot(m1)
```
```{r}
d %>% group_by(Geno) %>%
  summarise(mean = mean(CT_change, na.rm = TRUE),
            sd = sd(CT_change, na.rm = TRUE)) %>%
  kable(digits = 2)

d %>% group_by(Exercise) %>%
  summarise(mean = mean(CT_change, na.rm = TRUE),
            sd = sd(CT_change, na.rm = TRUE)) %>%
  kable(digits = 2)

d %>% group_by(Geno, Exercise) %>%
  summarise(mean = mean(CT_change, na.rm = TRUE),
            sd = sd(CT_change, na.rm = TRUE)) %>%
  kable(digits = 2)
```

#### Total movement

```{r}
ggplot(d, aes(x = Geno, colour = Exercise, y = TM_change)) +
  geom_boxplot(position = position_dodge(width = dw), alpha = 0.0) +
  geom_quasirandom(dodge.width = dw, size = 2) +
  scale_colour_manual(values = c("black", "red")) +
  ylab("Change total movement (cm)") +
  xlab("Genotype") +
  theme_bw()
```

```{r}
m1 <- lm(TM_post ~ TM_pre + Geno*Exercise, d)
anova(m1)

plot(m1)
```

```{r}
d %>% group_by(Geno) %>%
  summarise(mean = mean(TM_change, na.rm = TRUE),
            sd = sd(TM_change, na.rm = TRUE)) %>%
  kable(digits = 2)

d %>% group_by(Exercise) %>%
  summarise(mean = mean(TM_change, na.rm = TRUE),
            sd = sd(TM_change, na.rm = TRUE)) %>%
  kable(digits = 2)

d %>% group_by(Geno, Exercise) %>%
  summarise(mean = mean(TM_change, na.rm = TRUE),
            sd = sd(TM_change, na.rm = TRUE)) %>%
  kable(digits = 2)



```


### Grip strength

```{r}
gs <- select(d, ID, Geno, Exercise, HS_pre, HS_post, FS_pre, FS_post) %>%
  melt(id.vars = c("ID", "Geno", "Exercise")) %>%
  mutate(time = case_when(variable %in% c("HS_pre", "FS_pre") ~ "Pre",
                          variable %in% c("HS_post", "FS_post") ~ "Post"),
         time = factor(time, levels = c("Pre", "Post"))) %>%
  mutate(Limb = case_when(variable %in% c("HS_pre", "HS_post") ~ "Hindlimb",
                          variable %in% c("FS_pre", "FS_post") ~ "Forelimb"))

ggplot(gs, aes(x = Geno, colour = Exercise, y = value, shape = Limb)) +
  geom_boxplot(position = position_dodge(width = dw), alpha = 0) +
    geom_quasirandom(dodge.width = dw, size = 2, alpha = 0.5) +
  scale_colour_manual(values = c("black", "red")) +
  facet_wrap(~time) +
  ylab("Grip strength") +
  xlab("Genotype") +
  theme_bw()
```

Mixed effects models were fitted with a random intercept for animal id using the `lme4` package in R. The Genotype x Exercise x time 3-way interaction, and Genotype x Exercise  and Genotype x time interactions were tested, adjusting for limb (front or hind), and Exercise x time interaction.  

```{r}
m1 <- lmer(value ~ Limb + Geno*time + Exercise*time + Geno*Exercise*time + (1|ID), gs)
anova(m1)

gs %>% group_by(Geno, time) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE)) %>%
  kable(digits = 2)

gs %>% group_by(Geno, Exercise, time) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE)) %>%
  kable(digits = 2)
```

```{r, fig.cap = "Mean grip strength [95% CI]"}
library(plotrix)
gs0 <- gs %>% group_by(time, Exercise, Geno) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            se = std.error(value, na.rm = TRUE)) %>%
  mutate(lwr = mean - 1.96*se,
         upr = mean + 1.96*se)

ggplot(gs0, aes(x = time, colour = Geno, group = Geno, y = mean, ymin = lwr, ymax = upr)) +
  geom_point(size = 3, position = position_dodge(width = 0.2)) +
  geom_linerange(position = position_dodge(width = 0.2), size = 1) +
  geom_path(position = position_dodge(width = 0.2), linetype = "dashed", size = 1) +
  scale_colour_manual(values = c("black", "red")) +
  ylab("Grip strength /gm bw") +
  xlab("") +
  facet_wrap(~Exercise) +
  theme_bw()

```

### Foot print

```{r}
gs <- select(d, id, geno, exercise, fp_pre, fp_post) %>%
  melt(id.vars = c("id", "geno", "exercise")) %>%
  mutate(time = factor(variable, labels = c("Pre", "Post"))) %>%
  na.omit()

ggplot(gs, aes(x = geno, colour = exercise, y = value)) +
  geom_boxplot(position = position_dodge(width = 0.666)) +
    geom_quasirandom(dodge.width = 0.666) +
  scale_colour_manual(values = c("black", "red")) +
  facet_wrap(~time) +
  ylab("Foot print") +
  xlab("Genotype") +
  theme_bw()
```

Mixed effects models were fitted with a random intercept for animal id using the `lme4` package in R. All two-way interactions were tested.  

```{r}
m1 <- lmer(value ~ geno*time + exercise*time + geno*exercise + (1|id), gs)
anova(m1)

gs %>% group_by(geno, time) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE)) %>%
  kable()

gs %>% group_by(exercise, time) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE)) %>%
  kable()

gs %>% group_by(geno, exercise, time) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE)) %>%
  kable()
```
