---
title: "NFL behaviour"
author: "Bindoff, A."
output: html_document
---

`r Sys.time()`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## {.tabset}

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(knitr)
library(reshape2)
library(ggbeeswarm)
d <- read_csv("nfl_exercise.csv") %>%
  mutate(exercise = factor(exercise, levels = c(1, 0), labels = c("exercise", "no exercise")),
         geno = factor(geno, levels = c("WT", "NFL")))

```

### Open field

#### Centre time

```{r}
ggplot(d, aes(x = geno, colour = exercise, y = of_change_centre_time)) +
  geom_boxplot(position = position_dodge(width = 0.666)) +
  geom_quasirandom(dodge.width = 0.666) +
  scale_colour_manual(values = c("black", "red")) +
  ylab("Change time in centre (min? s?)") +
  xlab("Genotype") +
  theme_bw()
```

```{r}
m1 <- lm(of_change_centre_time ~ geno*exercise, d)
anova(m1)

d %>% group_by(geno) %>%
  summarise(mean = mean(of_change_centre_time, na.rm = TRUE),
            sd = sd(of_change_centre_time, na.rm = TRUE)) %>%
  kable()

d %>% group_by(exercise) %>%
  summarise(mean = mean(of_change_centre_time, na.rm = TRUE),
            sd = sd(of_change_centre_time, na.rm = TRUE)) %>%
  kable()

d %>% group_by(geno, exercise) %>%
  summarise(mean = mean(of_change_centre_time, na.rm = TRUE),
            sd = sd(of_change_centre_time, na.rm = TRUE)) %>%
  kable()
```

#### Total movement

```{r}
ggplot(d, aes(x = geno, colour = exercise, y = of_change_total_movement)) +
  geom_boxplot(position = position_dodge(width = 0.666)) +
  geom_quasirandom(dodge.width = 0.666) +
  scale_colour_manual(values = c("black", "red")) +
  ylab("Change total movement (cm? mm?)") +
  xlab("Genotype") +
  theme_bw()
```

```{r}
m1 <- lm(of_change_total_movement ~ geno*exercise, d)
anova(m1)

d %>% group_by(geno) %>%
  summarise(mean = mean(of_change_total_movement, na.rm = TRUE),
            sd = sd(of_change_total_movement, na.rm = TRUE)) %>%
  kable()

d %>% group_by(exercise) %>%
  summarise(mean = mean(of_change_total_movement, na.rm = TRUE),
            sd = sd(of_change_total_movement, na.rm = TRUE)) %>%
  kable()

d %>% group_by(geno, exercise) %>%
  summarise(mean = mean(of_change_total_movement, na.rm = TRUE),
            sd = sd(of_change_total_movement, na.rm = TRUE)) %>%
  kable()



```


### Grip strength

```{r}
gs <- select(d, id, geno, exercise, gs_pre, gs_post) %>%
  melt(id.vars = c("id", "geno", "exercise")) %>%
  mutate(time = factor(variable, labels = c("Pre", "Post"))) %>%
  na.omit()

ggplot(gs, aes(x = geno, colour = exercise, y = value)) +
  geom_boxplot(position = position_dodge(width = 0.666)) +
    geom_quasirandom(dodge.width = 0.666) +
  scale_colour_manual(values = c("black", "red")) +
  facet_wrap(~time) +
  ylab("Grip strength") +
  xlab("Genotype") +
  theme_bw()
```

Mixed effects models were fitted with a random intercept for animal id using the `lme4` package in R. All two-way interactions were tested and the three-way interaction for geno x exercise x time.  

```{r}
m1 <- lmer(value ~ geno*time + exercise*time + geno*exercise*time + (1|id), gs)
anova(m1)

gs %>% group_by(geno, time) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE)) %>%
  kable()

gs %>% group_by(exercise, time) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE)) %>%
  kable()

gs %>% group_by(geno, exercise, time) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE)) %>%
  kable()
```

Significant main effect of genotype and a significant time x exercise interaction.  We can plot this interaction more simply -   

```{r, fig.cap = "Mean $\\pm$ 1SE, interaction shown dashed"}
library(plotrix)
gs0 <- gs %>% group_by(exercise, time) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            se = std.error(value, na.rm = TRUE)) %>%
  mutate(lwr = mean - se,
         upr = mean + se)

ggplot(gs0, aes(x = time, colour = exercise, group = exercise, y = mean, ymin = lwr, ymax = upr)) +
  geom_point(size = 3, position = position_dodge(width = 0.2)) +
  geom_linerange(position = position_dodge(width = 0.2), size = 1) +
  geom_path(position = position_dodge(width = 0.2), linetype = "dashed", size = 1) +
  scale_colour_manual(values = c("black", "red")) +
  ylab("Grip strength") +
  xlab("Genotype") +
  theme_bw()

```

### Foot print

```{r}
gs <- select(d, id, geno, exercise, fp_pre, fp_post) %>%
  melt(id.vars = c("id", "geno", "exercise")) %>%
  mutate(time = factor(variable, labels = c("Pre", "Post"))) %>%
  na.omit()

ggplot(gs, aes(x = geno, colour = exercise, y = value)) +
  geom_boxplot(position = position_dodge(width = 0.666)) +
    geom_quasirandom(dodge.width = 0.666) +
  scale_colour_manual(values = c("black", "red")) +
  facet_wrap(~time) +
  ylab("Foot print") +
  xlab("Genotype") +
  theme_bw()
```

Mixed effects models were fitted with a random intercept for animal id using the `lme4` package in R. All two-way interactions were tested.  

```{r}
m1 <- lmer(value ~ geno*time + exercise*time + geno*exercise + (1|id), gs)
anova(m1)

gs %>% group_by(geno, time) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE)) %>%
  kable()

gs %>% group_by(exercise, time) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE)) %>%
  kable()

gs %>% group_by(geno, exercise, time) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE)) %>%
  kable()
```
